Pattern matching for\\ $(231,213)$-avoiding permutations

### 2015-02-03 article.tex
modification : in section 4 used factor instead of segment and used left most element instead of important  \n add : maccro for factor and left most element

@@

			For the proof and the algorithm we introduce a decomposition,
			we decompose the motif into
			adjacent element of same step in the stripe 
			(called such sequence segment),
			add the last element to the right most segment,
			and label them form right to left
			with increasing number.
			Let's call this decomposition $S$.
			For each segment, we call important 
			the leftmost element denoted by $IMP$.
			For example, given the motif $(12,11,10,1,2,3,4,9,8,5,6,7)$,
			$S(1)=(5,6,7)$, $IMP(1)=5$, 
			$S(2)=(9,8)$, $IMP(2)=9$, 
			$S(3)=(1,2,3,4)$, $IMP(3)=1$, 
			$S(4)=(12,11,10)$ and $IMP(4)=12$.\\

			\begin{proof}
			Given a ascent (descent) segment $S(s)$,
			if we want to
			match the suffix $s_ns_{s-1}...s_1$
			starting with an ascent (descent),
			and given that we know every
			match of $s_{s-1}...s_1$,
			the optimal way of matching
			$s_{s-1}...s_1$
			is to choose the match
			that minimize (maximize)
			the maximal (minimal)
			element of the match
			and that allow a match
			of $s_n$.			
			In other word if we want to match 
			$\pmotif[IMP(s):]$
			with $\ptext[i:]$
			and match $IMP(s)$
			with $i$.
			The optimal way
			of matching
			$\pmotif[IMP(s-1):]$
			is to choose the
			minimal (maximal) element $\ptext[i']$
			such that :
			(1) $\pmotif[IMP(s-1):]$ is matched with $\ptext[i':]$
			and $IMP(s-1)$ is matched with $e$.
			(2) The segment $S(s)$ is matched with $\ptext[i:i'-1]$
			and $IMP(s)$ is matched with $i$.
			(3) Every elements of the match of (2),
			is inferior (superior) to the minimal (maximal) element of the match of (1).
			Indeed when we will 
			match $s_{n+1}s_ns_{s-1}...s_1$
			more element will be available,
			because every matched 
			element of $s_{n+1}$
			must be superior (inferior)
			to every matched element
			of $s_{s-1}...s_1$.\\

			Consider the following problem :\\
			$LM(s,i)$ = the optimal match of 	
				$IMP(s-1)$ 
				if there exists a match of $\pmotif[IMP(s):]$ 
				with $\ptext[i:]$
				and $IMP(s)$ 
				is matched with element $i$.\\

@@

			\textbf{BASE :} \\
			$LM(1,i)= \left\{ 
					\begin{array}{l l}

						MIN_{i<i'}\{0\} \cup \{i' |\text{ $i'$ such that } & 
						\text{if $\stripe{\pmotif}{IMP(1)}=\ustep$}
						\\
						\text{ $LIS(i,i',\ptext[i']) \leq S(1)$}\}&\\

@@

						MAX_{i<i'}\{0\} \cup \{i' |\text{ $i'$ such that } &
						\text{if $\stripe{\pmotif}{IMP(1)}=\dstep$} \\
						\text{ $LDS(i,i',\ptext[i']) \leq S(1)$}\}&\\

@@

			\textbf{STEP :} \\
			$LM(s,i)= \left\{ 
					\begin{array}{l l}

						MIN \{0\} \cup  I(s,i) & 
						\text{if $\stripe{\pmotif}{IMP(s)}=\ustep$}\\

						MAX \{0\} \cup  D(s,i) & 
						\text{if $\stripe{\pmotif}{IMP(s)}=\ustep$}\\

@@

			With $I(s,i)$ ($D(s,i)$) the set of element such that
			if  $e \in I(s,i)$ then there exists 
			a match of $\pmotif[IMP(s-1):]$ in $\ptext[e:]$ and
			there is a match of $s$ in $\ptext[i:e-1]$
			with every element inferior (superior) to the match of  
			$\pmotif[IMP(s-1):]$ in $\ptext[e:]$.
			Formally we define $I(s,i)$ and $D(s,i)$ such as:\\

			$I(s,i)=\{i'|\text{$i<i'$ and $LM(s-1,i')\neq 0$}$ \\
			\indent \indent \indent and  $LIS(i,i'-1,IMP(s-1)) \leq S(s) \}$ \\

			$D(s,i)=\{i'|\text{$i<i'$ and $LM(s-1,i')\neq 0$}$ \\
			\indent \indent \indent and  $LDS(i,i'-1,IMP(s-1)) \leq S(s) \}$ \\

@@

			In the base case, 
			we are looking for a match for the first segment.
			Each segment is either ascent or descent.
			If the segment is an ascent (descent)
			we have to find an increasing (decreasing) sequence
			in the text of same size of longest that 
			the size of the segment,
			and to find the optimal solution 
			we must assure that the last element
			of the sequence is minimal (maximal).\\
			For the induction, it is the same idea
			except that we must assure that 
			every element of the sequence is superior 
			to the maximum element of the rest of the match
			or inferior to the minimal element of the rest of the match
			which is given by $LM$ of
			the previous segment.\\

			There exists a match of $\pmotif$ in $\ptext$ if and only if
			there exists a $LM(n,i)$ for $1 \leq i \leq |\ptext|$,
			with $n$ the number of segment in $\pmotif$.
			Moreover the basic case can be computed in $O(|\ptext|^2)$
			and the induction on $O(|\pmotif|*|\ptext|^2)$.
			\end{proof}

>>>>

%first
\DeclareMathOperator{\firsta}{L}
\newcommand{\first}[1]{\firsta({#1})}

\DeclareMathOperator{\factora}{F}
\newcommand{\factor}[1]{\factora({#1})}

@@

For the proof and the algorithm we introduce a decomposition : we decompose the stripe of the motif into factor of same letter, and label those factor from right to left. We also need to add the last index into the first factor. For each factor, we denote the left most element by $\firsta$. For example, given the motif $981237654$, $\factor{1} =7654$, $\first{1}=7$, $\factor{2}=123$, $\first{2}=1$, $\factor{3}=98$ and $\first{9}$. \\

\begin{proof}
Given an ascent (descent) factor $\factor{i}$ of a motif $\pmotif$, if we want to match the suffix of $\pmotif$ starting by $\factor{i}$, $\factor{i}\factor{i-1}\cdots\factor{1}$ and given that we know every match of $\factor{i-1}\cdots\factor{1}$, is to choose the match of $\factor{i-1}\cdots\factor{1}$ that minimize (maximize) the maximal (minimal) element of the match of $\factor{i-1}\cdots\factor{1}$ and that allow a match of factor $\factor{i}$. In other word if we want to match $\pmotif[\first{i}:]$ with $\ptext[j:]$ such that $\first{i}$ is matched with element at index $j$ in $\ptext$. The optimal way of matching $\pmotif[\first{i-1}:]$ is to choose the minimal (maximal) element $j'$ of $\ptext[j:]$ such that : (1) $\pmotif[\first{i-1}:]$ is matched with $\ptext[j':]$ and element $\first{i-1}$ is matched with element at index $j'$, (2) the factor $\factor{i}$ is matched with $\ptext[j,j'-1]$ and element $\first{i}$ is matched with element at index $j$, (3) Every elements of the match of (2) if inferior (superior) to the minimal (maximal) element of the match of (1). Indeed when we will match $\factor{i+1}\factor{i}$, $\factor{i}\factor{i-1}\cdots\factor{1}$, every elements of $\factor{i+1}$ must be matched to element superior (inferior) to every element matched element of $\factor{i}\factor{i-1}\cdots\factor{1}$, particularly its must be superior (inferior) to the maximal (minimal) element of the match of $\factor{i}\factor{i-1}\cdots\factor{1}$. Remark that the maximal (minimal) element of $\factor{i}\factor{i-1}\cdots\factor{1}$ is $\first{i-1}$.\\

Consider the following problem :\\
$LM(i,j)$ = the optimal match of $\first{i-1}$	
	if there exists a match of $\pmotif[\first{i}:]$ 
	with $\ptext[j:]$
	and element $\first{i}$ 
	is matched with element at index $j$.\\

@@

\textbf{BASE :} \\
$LM(1,j)= \left\{ 
	\begin{array}{l l}

		MIN_{j<j'}\{0\} \cup \{\ptext[j'] |\text{ $j'$ such that \indent } & 
		\text{if $\stripe{\pmotif}{\first{1}}=\ustep$}\\
		\text{ $LIS(j,j',\ptext[j']) \leq \factor{1}$}\}&\\

@@

		MAX_{j<j'}\{0\} \cup \{\ptext[j'] |\text{ $j'$ such that \indent } &
		\text{if $\stripe{\pmotif}{\first{1}}=\dstep$} \\
		\text{ $LDS(j,j',\ptext[j']) \leq \factor{1}$}\}&\\

@@

\textbf{STEP :} \\
$LM(i,j)= \left\{ 
	\begin{array}{l l}

		MIN \{0\} \cup  I(i,j) & 
		\text{if $\stripe{\pmotif}{\first{i}}=\ustep$}\\

		MAX \{0\} \cup  D(i,j) & 
		\text{if $\stripe{\pmotif}{\first{i}}=\dstep$}\\

@@

With $I(i,j)$ ($D(i,j)$) the set of element such that
if  $j' \in I(i,j)$ then there exists 
a match of $\pmotif[\first{i-1}:]$ in $\ptext[j':]$ and
the element $\first{i-1}$ is matched with element at index $j'$ and
there is a match of $\factor{i}$ in $\ptext[j:j'-1]$ and
the element $\first{i}$ is matched with element at index $j$ and every element is inferior (superior) to the match of  
$\pmotif[\first{i-1}:]$ in $\ptext[j':]$.
Formally we define $I(s,i)$ and $D(s,i)$ such as:\\

$I(i,j)=\{j'|\text{$j<j'$ and $LM(i-1,j')\neq 0$}$ \\
\indent \indent \indent and  $LIS(j,j'-1,\first{i-1}) \leq |\factor{i}| \}$ \\

$D(i,j)=\{j'|\text{$j<j'$ and $LM(i-1,j')\neq 0$}$ \\
\indent \indent \indent and  $LDS(j,j'-1,\first{i-1}) \leq |\factor{i}| \}$ \\

@@

In the base case, 
we are looking for a match for the first factor.
Each factor is either ascent or descent.
If the factor is an ascent (descent)
we have to find an increasing (decreasing) subsequence
in the text of same size or longest that 
the size of the factor,
and to find the optimal solution 
we must assure that the last element
of the sequence is minimal (maximal).\\
For the induction, it is the same idea
except that we must assure that 
every element of the subsequence is superior 
to the maximum element of the rest of the match
ie inferior to the minimal (maximal) element of the rest of the match
which is given by $LM$ of
the previous factor.\\

There exists a match of $\pmotif$ in $\ptext$ if and only if
there exists a $LM(n,i)\neq 0$ for $1 \leq i \leq |\ptext|$,
with $n$ the number of factor in $\pmotif$.
Moreover the basic case can be computed in $O(|\ptext|^2)$
and the induction on $O(|\pmotif|*|\ptext|^2)$.
\end{proof}
